---
title: "solar cycle paper"
author: "Richard J. Telford"
date: "June 30, 2017"
output: html_document
bibliography: solarCycles.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, cache.lazy=TRUE)
knitr::read_chunk("ringwidths.r")
```

```{r loadPackages, cache=FALSE, message=FALSE}
```
```{r importData, fig.show='hide'}
```

## Abstract


**Keywords:** cyclicity, Schwabe cycle, solar cycle, tree-rings

## Introduction

Variability in solar output. TSI + enhanced variability in UV 
Instrumental evidence of effect of solar variability- regional or hemispheric
single stations??

Proxy data inevitably noisy. Long history of attempts to detect solar variabilty on proxy data. Focus on 11-year Schwabe cycle. Examples: tree rings (south american, Permian etc), varves (xxx). ice out dates

Problems: publication bias, autocorration, spectral analysis  - assumption of white noise background, or AR1 process etc - broad definitions - multiple testing.

Solution: systematic analysis of tree ring data for evidence of solar signals in time and frequency domain.

## Methods

selection of ring-width series. From St George et al. synthesis of N hemisphere records. Signal strength, Span period `r time_limits$min`--`r time_limits$max`. Former cutoff - more reliable sunspot data, latter cutoff, prevent dropout of series, avoid period with greatest human influence on climate and CO2 concentrations (N-loading).
Species

We use yearly sunspot data from [#cite] as this is the longest annual-resolution proxy of solar activity.

Spectral analysis

The spectra are summarised by finding the proportion that are statistically significanct  (p > 0.1) for each frequency. If the trees are sensitive to solar variability, there should be an excess of spectra with a significant cyclicity corresponding to the Scwabe cycle. 

Regression analyses

Sensitivity tests

`r gsub("(\\1) \\(.*", "\\1", R.version$version.string)` [@RCoreTeam] was used for all analyses.  


## Results

`r gsub("([[:alpha:]])(.+)", "\\U\\1\\L\\2", english::as.english(length(rwl)), perl = TRUE)` ringwidth series met the inclusion criteria (Fig 1). 

```{r map1, fig.height = 3, fig.cap="Figure 1. Location of ringwidth series that met the inclusion criteria."}
```

```{r frequency, fig.show = "hide", results = "hide"}
```

Over the time period studied the sunspots record has a strong spectral peak at 1/`r 
round(1/ssf$freq[which.max(ssf$spec)], 2)` years^-1^ (Fig. 2a). Some of the ringwidth series have a spectral peak at this frequency (Fig. 2b), but the less than 10% of the series have a spectral peak coincident with the Schwabe cycle (Fig. 2c). These series are geographically and taxonomically widespread.

```{r fig2, fig.keep = "last", fig.cap = "Figure 2. a) Sunspot spectrum, b) spectrum of example ringwidth series, c) percent of ringwidth series with significant periodicity (p < 0.1) at each frequency. Dashed line at 10% indicates the expected percentage."}
```



## Discussion

Hale cycle

## Conclusions

## References